Tree height and hydraulic traits shape growth responses across droughts in a temperate broadleaf forest

### 2020-02-20 manuscript/climate_sensitivity_draft.Rmd
minor updates

While some canopy positions undoubtedly shifted in the # years between the 1966 drought and our observations in 2018, change was relatively slow, and bias would more likely result in type x error with regards to testing our hypothesis.

@@

Our analysis indicates that tree height has a stronger influence on drought response than does canopy position (Tables 1,4,5). This is consistent with, and reinforces, previous findings that biophysical constraints make it impossible for trees to efficiently transport water to great heights and simultaneously maintain strong resistance and resilience to drought-induced embolism [@olson_plant_2018; @couvreur_water_2018; @roskilly_conflicting_2019]. However, this result must be interpeted with some caution, given that colinearity between the two variables (Fig. **2d**) makes it impossible to confidently partition causality. Taller trees are more likely to be in dominant canopy positions (Fig. **2d**) and, largely as a consequence of their position relative to others, face different microenvironments (Fig. **2a-b**). Even under non-drought conditions, evaporative demand and maximum leaf teamperatures increase with tree height [@smith_temperature_1977; @bretfeld_plant_2018; @kunert_revised_2017], and such conditions would incur extra stress during drought, when solar radiation tends to be higher and less water is available for evaporative cooling of the leaves. However, some decoupling between height and canopy position is introduced by the configuration of neighboring trees (Fig. 2d) [@muller-landau_testing_2006], and height was an overall stronger predictor of drought response than crown position (Tables 1,4,5).

>>>>

While some canopy positions undoubtedly shifted in the 52 years between the 1966 drought and our observations in 2018, change was relatively slow, and bias would more likely result in type x error with regards to testing our hypothesis.

@@

Our analysis indicates that tree height has a stronger influence on drought response than does canopy position (Tables 1,4,5). This is consistent with, and reinforces, previous findings that biophysical constraints make it impossible for trees to efficiently transport water to great heights and simultaneously maintain strong resistance and resilience to drought-induced embolism [@olson_plant_2018; @couvreur_water_2018; @roskilly_conflicting_2019]. However, this result must be interpeted with some caution, given that colinearity between the two variables (Fig. 2d) makes it impossible to confidently partition causality. Taller trees are more likely to be in dominant canopy positions (Fig. 2d) and, largely as a consequence of their position relative to others, face different microenvironments (Fig. 2a-b). Even under non-drought conditions, evaporative demand and maximum leaf teamperatures increase with tree height [@smith_temperature_1977; @bretfeld_plant_2018; @kunert_revised_2017], and such conditions would incur extra stress during drought, when solar radiation tends to be higher and less water is available for evaporative cooling of the leaves. However, some decoupling between height and canopy position is introduced by the configuration of neighboring trees (Fig. 2d) [@muller-landau_testing_2006], and height was an overall stronger predictor of drought response than crown position (Tables 1,4,5).
